---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Jackson Menoher"
output:
  html_document:
    df_print: paged
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("states.csv")
```

### Exercise 1
```{r}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```


### Exercise 2
```{r}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```

### Exercise 3
```{r}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak
```


### Exercise 4
```{r}
nrow(dn_lq_ak)
```
Six rows in the full joined dataframe. 

```{r}
colnames(dn_lq_ak)
```
The variable names are shown above. 

### Exercise 5
```{r}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```


### Exercise 6
```{r}
dn_lq_ak <- dn_lq_ak %>% 
  mutate(distance = 
           haversine(longitude.x, latitude.x, longitude.y, latitude.y,round=4))
```

### Exercise 7
```{r}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```

### Exercise 8
```{r}
hist(dn_lq_ak_mindist$closest)
```
```{r}
summary(dn_lq_ak_mindist)
```
### Exercise 9
```{r}
dn_nc <- dennys %>%
  filter(state == "NC")

lq_nc <- laquinta %>%
  filter(state == "NC")

dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")

dn_lq_nc <- dn_lq_nc %>% 
  mutate(distance = haversine(longitude.x, latitude.x,
                              longitude.y, latitude.y))

dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))

hist(dn_lq_nc_mindist$closest)

summary(dn_lq_nc_mindist)

```

### Exercise 10
```{r}
dn_tx <- dennys %>%
  filter(state == "TX")

lq_tx <- laquinta %>%
  filter(state == "TX")

dn_lq_tx <- full_join(dn_tx, lq_tx, by = "state")

dn_lq_tx <- dn_lq_tx %>% 
  mutate(distance = haversine(longitude.x, latitude.x,
                              longitude.y, latitude.y))

dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))

hist(dn_lq_tx_mindist$closest)

summary(dn_lq_tx_mindist)
```

### Exercise 11
```{r}
dn_ny <- dennys %>%
  filter(state == "NY")

lq_ny <- laquinta %>%
  filter(state == "NY")

dn_lq_ny <- full_join(dn_ny, lq_ny, by = "state")

dn_lq_ny <- dn_lq_ny %>% 
  mutate(distance = haversine(longitude.x, latitude.x,
                              longitude.y, latitude.y))

dn_lq_ny_mindist <- dn_lq_ny %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))

hist(dn_lq_ny_mindist$closest)

print(mean(dn_lq_ny_mindist$closest))
print(sd(dn_lq_ny_mindist$closest))
```

### Exercise 12
Among the states AK, TX, and NY, Mitch Hedberg's joke is most likely to hold up in Alaska (AK) being that Alaska has the smallest average distance between the Denny's and LaQuinta's establishments. 
